---
title: "Analysis 2: Conservation effectiveness since 1963"
output: html_document
---

```{r setup, include=FALSE}

# set you working directory 
setwd(my_wd <- 'path/to/your/data/dir/') 
in_path <- paste0(my_wd,"input/")
out_path <- paste0(my_wd,"output/")

library(dplyr)
library(reshape2)
```

# 1 Changes in the number and area of protected areas since 1963

```{r}

out_csv <- paste0(out_path, "result_2_conservation_effectiveness_since1963/Plot_number_and_area_PAs.csv")
if(!file.exists(out_csv)) {
  
  path_prop <- paste0(out_path,"result_1_conservation_gaps/prop_covered_habitat_by_species.csv")
  path_sp <- read.csv(path_prop)$gbif_name
  
  path_PAs <- paste0(in_path, "PAs_coverage/PAs_by_species_based_AOH30m_aggregated.rds")
  df_PAs <- readRDS(path_PAs) %>% 
    select(-all_of(path_sp)) %>% 
    group_by(PAs_ID) %>%
    summarize(across(c(No_pixel, No_pixel_PAs), sum),
              across(-c(No_pixel, No_pixel_PAs), ~ first(.[!is.na(.)]))) %>% 
    select(-c(PAs_value, No_pixel, NR_ID, NP_ID)) %>% 
    mutate(count = 1) %>% 
    setNames(gsub("_PAs", "", colnames(.)))
  
  df_panda_PAs <- df_PAs %>% 
    filter(type==1) %>% 
    group_by(Year) %>%
    summarize(across(c(No_pixel, count), sum),
              across(-c(No_pixel, count), ~ first(.[!is.na(.)]))) %>% 
    select(Year, No_pixel, count) %>% 
    setNames(c("Year", "Pixel_Panda", "Number_Panda")) %>% 
    mutate(Total_Pixel_Panda = cumsum(Pixel_Panda),
           Total_Number_Panda = cumsum(Number_Panda),
           Total_Area_Panda = Total_Pixel_Panda*(30/1000)^2)
  
  df_plant_PAs <- df_PAs %>% 
    filter(type==-1) %>% 
    group_by(Year) %>%
    summarize(across(c(No_pixel, count), sum),
              across(-c(No_pixel, count), ~ first(.[!is.na(.)]))) %>% 
    select(Year, No_pixel, count) %>% 
    setNames(c("Year", "Pixel_Plant", "Number_Plant")) %>% 
    mutate(Total_Pixel_Plant = cumsum(Pixel_Plant),
           Total_Number_Plant = cumsum(Number_Plant),
           Total_Area_Plant = Total_Pixel_Plant*(30/1000)^2)
  
  df_other_PAs <- df_PAs %>% 
    filter(is.na(type)) %>%
    mutate(type=0) %>% 
    group_by(Year) %>%
    summarize(across(c(No_pixel, count), sum),
              across(-c(No_pixel, count), ~ first(.[!is.na(.)]))) %>% 
    select(Year, No_pixel, count) %>% 
    setNames(c("Year", "Pixel_Other", "Number_Other")) %>% 
    mutate(Total_Pixel_Other = cumsum(Pixel_Other),
           Total_Number_Other = cumsum(Number_Other),
           Total_Area_Other = Total_Pixel_Other*(30/1000)^2)
  
  df_panda_plant <- merge(df_panda_PAs, df_plant_PAs, by= "Year", all=T)
  df_panda_plant_other <- merge(df_panda_plant, df_other_PAs,by="Year", all=T)
  
  df_all_PAs <- df_PAs %>% 
    group_by(Year) %>%
    summarize(across(c(No_pixel, count), sum),
              across(-c(No_pixel, count), ~ first(.[!is.na(.)]))) %>% 
    select(Year, No_pixel, count) %>% 
    setNames(c("Year", "Pixel_All", "Number_All")) %>% 
    mutate(Total_Pixel_All = cumsum(Pixel_All),
           Total_Number_All = cumsum(Number_All),
           Total_Area_All = Total_Pixel_All*(30/1000)^2)
  
  df_PAs_merged <- merge(df_all_PAs, df_panda_plant_other, by="Year", all=T) %>% 
    select(Year, matches("Total_Number"), matches("Total_Area")) %>% 
    select(Year, matches("_All"), matches("_Panda"), matches("_Plant"), matches("_Other"))
  
  tmp_types <- c("Panda","Plant","Other")
  for (tmp_type in tmp_types) {
    rows <- which(is.na(df_PAs_merged[,paste0("Total_Number_",tmp_type)]))
    cols <- grepl(tmp_type, colnames(df_PAs_merged))
    if(length(rows)!=0){
      for (row in rows) {
        if (row==1) {
          df_PAs_merged[1,cols] <- 0
        } else {
          df_PAs_merged[row,cols] <- df_PAs_merged[row-1,cols]
        }
      }
    }
  }
  write.csv(df_PAs_merged, out_csv)
}

```


# 2 Conservation effectiveness in number of covered species since 1963

## 2.1 Add columns to group PAs by their establishment year into specific time intervals

```{r}

start_years <- 1963
end_years <- 2023
interval_years <- c(1,30)
suffix_name <- gsub("1years","1year",paste0("every_",interval_years,"years"))

out_rds <- paste0(out_path,"result_2_conservation_effectiveness_since1963/PAs_by_species_grouped.rds")
if (!file.exists(out_rds)) {
  
  path_PAs <- paste0(in_path, "PAs_coverage/PAs_by_species_based_AOH30m_aggregated.rds")
  df_PAs <- readRDS(path_PAs)
  for (i in length(interval_years):1) {
    if (interval_years[i]!=1) {
      time_nodes <- seq(start_years, end_years, by = interval_years[i])
      df_PAs <- df_PAs %>%
        mutate(!!suffix_name[i] := as.numeric(as.character(cut(Year, breaks = c(floor(start_years/10)*10,time_nodes), labels = time_nodes)))) %>%
        select(!!suffix_name[i], everything())
    }
    if (interval_years[i]==1) {
      time_nodes <- c(unique(df_PAs$Year[(order(df_PAs$Year))]), end_years)
      df_PAs <- df_PAs %>%
        mutate(!!suffix_name[i] := 
                 as.numeric(as.character(cut(Year, breaks = c(time_nodes, ceiling(end_years/10)*10), labels = time_nodes,right = FALSE)))) %>%
        select(!!suffix_name[i], everything()) %>%
        select(Year, everything())
    }
  }
  saveRDS(df_PAs, out_rds)
}

```

## 2.2 Group the PAs according to their conservation targets

```{r}

interval_years <- c(1,30)
suffix_name <- gsub("1years","1year",paste0("every_",interval_years,"years"))

for (i in 1:length(interval_years)) {
  out_rds_all <- paste0(out_path,"result_2_conservation_effectiveness_since1963/PAs_by_species_grouped_All_PAs_",suffix_name[i],".rds")
  out_rds_panda <- gsub("_All","_Panda",out_rds_all)
  out_rds_plant <- gsub("_All","_Plant",out_rds_all)
  out_rds_other <- gsub("_All","_Other",out_rds_all)
  out_rds <- c(out_rds_all, out_rds_panda, out_rds_plant, out_rds_other)
  if (sum(!file.exists(out_rds))!=0) {
    if (!file.exists(out_rds_all)) {
      path_PAs <- paste0(out_path,"result_2_conservation_effectiveness_since1963/PAs_by_species_grouped.rds")
      df_PAs <- readRDS(path_PAs)
      df_all <- aggregate(df_PAs, by=list(df_PAs[,suffix_name[i]]), FUN=sum, na.rm=T) %>% 
        mutate(Year = Group.1) %>% 
        select(-matches("_ID|PAs_|every_|Group.1|^type$", ignore.case = F))
      saveRDS(df_all, out_rds_all)
    }
    if(!file.exists(out_rds_panda)){
      df_panda <- df_PAs %>% 
        filter(type == 1)
      df_panda <- aggregate(df_panda, by=list(df_panda[,suffix_name[i]]), FUN=sum, na.rm=T)  %>% 
        mutate(Year = Group.1) %>% 
        select(-matches("_ID|PAs_|every_|Group.1|^type$", ignore.case = F))
      saveRDS(df_panda, out_rds_panda)
    }
    if(!file.exists(out_rds_plant)){
      df_plant <- df_PAs %>% 
        filter(type == -1)
      df_plant <- aggregate(df_plant, by=list(df_plant[,suffix_name[i]]), FUN=sum, na.rm=T)  %>% 
        mutate(Year = Group.1) %>% 
        select(-matches("_ID|PAs_|every_|Group.1|^type$", ignore.case = F))
      saveRDS(df_plant, out_rds_plant)
    }            
    if(!file.exists(out_rds_other)){
      df_other <- df_PAs %>% 
        filter(is.na(type))
      df_other <- aggregate(df_other, by=list(df_other[,suffix_name[i]]), FUN=sum, na.rm=T)  %>% 
        mutate(Year = Group.1) %>% 
        select(-matches("_ID|PAs_|every_|Group.1|^type$", ignore.case = F))
      saveRDS(df_other, out_rds_other)
    }  
  } 
}

        
```

## 2.3 Effectiveness of PAs in species coverage by conservation target within specific time intervals

```{r}

suffix_name <- "every_1year"
PA_types <- c("All_PAs", "Panda_PAs", "Plant_PAs", "Other_PAs")

for (tmp_PAtype in PA_types) {
  out_rds <- paste0(out_path,"result_2_conservation_effectiveness_since1963/Measure1_",tmp_PAtype,"_",suffix_name,".rds")
  if (sum(!file.exists(out_rds))!=0) {
    path_PAs <- paste0(out_path,"result_2_conservation_effectiveness_since1963/PAs_by_species_grouped_",tmp_PAtype,"_",suffix_name,".rds")
    tmp_rds <- readRDS(path_PAs)
    tmp_rds_binary <- tmp_rds %>% 
      vegan::decostand(method = "pa") %>% 
      mutate(Year = tmp_rds$Year,
             No_pixel= tmp_rds$No_pixel_PAs)
    
    tmp_summary <- data.frame(matrix(NA, nrow=nrow(tmp_rds_binary), ncol=6))
    colnames(tmp_summary) <- c("Year", "measure1", "increase_measure1",
                               "accumulate_measure1","area_km2", "accumulate_area")     
  
    tmp_summary <- data.frame(matrix(NA, nrow = nrow(tmp_rds_binary), ncol = 6)) %>%
      setNames(c("Year", "measure1", "increase_measure1", "accumulate_measure1", 
                 "area_km2", "accumulate_area")) %>%
      mutate(Year = tmp_rds_binary$Year,
             area_km2 = (tmp_rds_binary$No_pixel) * (30/1000)^2,
             accumulate_area = cumsum(area_km2))
    tmp_rds_binary <- tmp_rds_binary %>% 
      select(-matches("Year|No_pixel", ignore.case = F))
    
    for (s in 1:nrow(tmp_summary)) {
      tmp_summary$measure1[s] <- sum(colSums(tmp_rds_binary[s,]))
      tmp_summary$accumulate_measure1[s] <- sum(colSums(tmp_rds_binary[1:s,])!=0)
      if (s==1) {tmp_summary$increase_measure1[s] <- sum(colSums(tmp_rds_binary[1:s,])!=0)-0} 
      if (s!=1) {tmp_summary$increase_measure1[s] <-
        sum(colSums(tmp_rds_binary[1:s,])!=0)-sum(colSums(tmp_rds_binary[1:(s-1),])!=0)}
    }
    colnames(tmp_summary) <- paste0(colnames(tmp_summary),"_", tmp_PAtype)
    saveRDS(tmp_summary, out_rds)
  }
}

```

## 2.4 Merge different types of PAs for measure 1 into one table within specific intervals

```{r}

suffix_name <- "every_1year"
PA_types <- c("All_PAs", "Panda_PAs", "Plant_PAs", "Other_PAs")

out_rds <- paste0(out_path,"result_2_conservation_effectiveness_since1963/Measure1_",length(PA_types),"Types_PAs_",suffix_name,"_merged.rds")

if (sum(!file.exists(out_rds))!=0) {
  tmp_ff_rds <- list.files(paste0(out_path,"result_2_conservation_effectiveness_since1963/"),             paste0("Measure1.*",suffix_name,".rds$"),full.names = T)

  df_merged <- readRDS(tmp_ff_rds[1]) %>% 
    select(matches("Year")) %>% 
    setNames("Year") %>% 
    data.frame()
    
  for (tmp_PAtype in PA_types) {
    tmp_ff <- tmp_ff_rds[grepl(tmp_PAtype,tmp_ff_rds)]
    tmp_rds <- readRDS(tmp_ff) %>%
      select(matches(c("Year","accumulate"))) %>%
      rename(Year = matches("Year")) %>% 
      rename_all(~ gsub("_PAs|accumulate_", "", .))  %>% 
      data.frame()
    df_merged <- merge(df_merged, tmp_rds, by="Year", all=T) %>% data.frame
  }
  saveRDS(df_merged, out_rds)
}
  
```

## 2.5 Prepare the datasets for plotting the effectiveness of measure 1

```{r}

suffix_name <- "every_1year"
PA_types <- c("All_PAs", "Panda_PAs", "Plant_PAs", "Other_PAs")

out_csv <- paste0(out_path,"result_2_conservation_effectiveness_since1963/Plot_Measure1_",length(PA_types),"Types_PAs_",suffix_name,".csv")
out_csv_melt <- gsub("\\.csv","\\_melt.csv",out_csv)
out_all <- c(out_csv, out_csv_melt)

if(sum(!file.exists(out_all)!=0)) {
  
  path_PAs <- paste0(out_path,"result_2_conservation_effectiveness_since1963/Measure1_",length(PA_types),"Types_PAs_",suffix_name,"_merged.rds")
  df_all <- cbind(
    readRDS(path_PAs) %>% 
      select(matches(c("Year","All"))) %>% 
      select(matches(c("Year","measure1"))) %>%
      rename_all(~ gsub("measure1_All_", "", .)) %>%
      melt(id.vars = "Year", variable.name = "groups", value.name = "all_measure1"),
    readRDS(path_PAs) %>% 
      select(matches(c("Year","All"))) %>%
      select(matches(c("Year","area"))) %>%
      rename_all(~ gsub("area_All_", "", .)) %>%
      melt(id.vars = "Year", variable.name = "groups_area", value.name = "all_area")) %>% 
    data.frame()
  if(table(as.character(df_all$Year)==as.character(df_all$Year.1))==nrow(df_all)){df_all$Year.1 <- NULL}
  if(table(as.character(df_all$groups)==as.character(df_all$groups_area))==nrow(df_all)){df_all$groups_area <- NULL}
  
  df_panda <- cbind(
    readRDS(path_PAs) %>% 
      select(matches(c("Year","Panda"))) %>% 
      select(matches(c("Year","measure1"))) %>%
      rename_all(~ gsub("richness_Panda_", "", .)) %>%
      melt(id.vars = "Year", variable.name = "groups", value.name = "panda_measure1"),
    readRDS(path_PAs) %>% 
      select(matches(c("Year","Panda"))) %>% 
      select(matches(c("Year","area"))) %>%
      rename_all(~ gsub("area_Panda_", "", .)) %>%
      melt(id.vars = "Year", variable.name = "groups_area", value.name = "panda_area")) %>% 
    data.frame()

  if(table(as.character(df_panda$Year)==as.character(df_panda$Year.1))==nrow(df_panda)){df_panda$Year.1 <- NULL}
  if(table(as.character(df_panda$groups)==as.character(df_panda$groups_area))==nrow(df_panda)){df_panda$groups_area <- NULL}

  df_plant <- cbind(
    readRDS(path_PAs) %>% 
      select(matches(c("Year","Plant"))) %>% 
      select(matches(c("Year","measure1"))) %>%
      rename_all(~ gsub("measure1_Plant_", "", .)) %>%
      melt(id.vars = "Year", variable.name = "groups", value.name = "plant_measure1"),
    readRDS(path_PAs) %>% 
      select(matches(c("Year","Plant"))) %>% 
      select(matches(c("Year","area"))) %>%
      rename_all(~ gsub("area_Plant_", "", .)) %>%
      melt(id.vars = "Year", variable.name = "groups_area", value.name = "plant_area")) %>% 
    data.frame()

  if(table(as.character(df_plant$Year)==as.character(df_plant$Year.1))==nrow(df_plant)){df_plant$Year.1 <- NULL}
  if(table(as.character(df_plant$groups)==as.character(df_plant$groups_area))==nrow(df_plant)){df_plant$groups_area <- NULL}            

  df_other <- cbind(
    readRDS(path_PAs) %>% 
      select(matches(c("Year","Other"))) %>% 
      select(matches(c("Year","measure1"))) %>%
      rename_all(~ gsub("measure1_Other_", "", .)) %>%
      melt(id.vars = "Year", variable.name = "groups", value.name = "other_measure1"),
    readRDS(path_PAs) %>% 
      select(matches(c("Year","Other"))) %>% 
      select(matches(c("Year","area"))) %>%
      rename_all(~ gsub("area_Other_", "", .)) %>%
      melt(id.vars = "Year", variable.name = "groups_area", value.name = "other_area")) %>% 
    data.frame()

  if(table(as.character(df_other$Year)==as.character(df_other$Year.1))==nrow(df_other)){df_other$Year.1 <- NULL}
  if(table(as.character(df_other$groups)==as.character(df_other$groups_area))==nrow(df_other)){df_other$groups_area <- NULL}
  
  df_combine <- cbind(df_all, df_panda, df_plant, df_other) %>% data.frame()
  
  if(table(as.character(df_combine$Year)==as.character(df_combine$Year.1))==nrow(df_combine)){df_combine$Year.1 <- NULL}
  if(table(as.character(df_combine$groups)==as.character(df_combine$groups.1))==nrow(df_combine)){df_combine$groups.1 <- NULL}
  if(table(as.character(df_combine$Year)==as.character(df_combine$Year.2))==nrow(df_combine)){df_combine$Year.2 <- NULL}
  if(table(as.character(df_combine$groups)==as.character(df_combine$groups.2))==nrow(df_combine)){df_combine$groups.2 <- NULL}            
  if(table(as.character(df_combine$Year)==as.character(df_combine$Year.3))==nrow(df_combine)){df_combine$Year.3 <- NULL}
  if(table(as.character(df_combine$groups)==as.character(df_combine$groups.3))==nrow(df_combine)){df_combine$groups.3 <- NULL}

  
  tmp_types <- c("panda","plant","other")
  for (tmp_type in tmp_types) {
    rows <- which(is.na(df_combine[,paste0(tmp_type,"_measure1")]))
    cols <- grepl(tmp_type, colnames(df_combine))
    if(length(rows)!=0){
      for (row in rows) {
        if(row==1){
          df_combine[row,cols] <- 0
        }else{
          df_combine[row,cols] <- df_combine[row-1,cols]
        }
      }
    }
  }

  
  if(!file.exists(out_csv)) {
    write.csv(df_combine, out_csv)
  }
  if(!file.exists(out_csv_melt)) {
    df_combine_measure1 <- df_combine %>% 
      select(matches(c("Year","groups","measure1"))) %>%
      mutate(Year_groups = paste0(Year, "_", groups)) %>% 
      select(matches("_")) %>% 
      melt(id.vars = "Year_groups", variable.name = "type", value.name = "measure1") %>% 
      mutate(Year = as.numeric(unlist(lapply(strsplit(Year_groups, "_"), "[[", 1))),
             Year_groups = NULL,
             type = gsub("_measure1", "", type),
             group_datasets = paste0(type, "_PAs")) %>% 
      select(c("Year", "group_datasets", "measure1"))
    
    write.csv(df_combine_measure1, out_csv_melt)
  }
}

```


# 3 Conservation effectiveness in proportion of covered habitat since 1963

## 3.1 Calculate the proportion of covered habitats per species

```{r Prop_PAs}

suffix_name <- "every_30years"
PA_types <- c("All_PAs", "Panda_PAs", "Plant_PAs", "Other_PAs")

for (tmp_PAtype in PA_types) {
  out_rds <- paste0(out_path,"result_2_conservation_effectiveness_since1963/Measure2_",tmp_PAtype,"_",suffix_name,".rds")
  if(!file.exists(out_rds)){
    
    path_SA <- paste0(in_path,"pixels_by_species/pixels_by_species_AOH30m_aggregated.rds")
    path_PA <- paste0(out_path,"result_2_conservation_effectiveness_since1963/PAs_by_species_grouped_",tmp_PAtype,"_",suffix_name,".rds")
    
    tmp_PAs <- readRDS(path_PA) %>% select(-No_pixel_PAs)
    tmp_SA <- readRDS(path_SA) %>% colSums(.);gc()
    
    tmp_PAs_SA <- tmp_PAs %>% 
      select(-c("Year", "No_pixel"))
    for (z in 2:nrow(tmp_PAs)) {
      tmp_PAs_sel <- tmp_PAs[1:z,] %>% select(-c("Year", "No_pixel"))
      tmp_PAs_SA[z,] <- colSums(tmp_PAs_sel)
    }
    tmp_PAs_SA <- tmp_PAs_SA %>%
      t() %>%
      data.frame() %>%
      setNames(paste0("Year_", tmp_PAs$Year)) %>%
      mutate(sum_pixels= rowSums(.)) / tmp_SA
    
    saveRDS(tmp_PAs_SA, out_rds)
  }
}


```

## 3.2 Frequency of habitat coverage proportions

```{r}

my_break <- 0.2
suffix_name <- "every_30years"
PA_types <- c("All_PAs", "Panda_PAs", "Plant_PAs", "Other_PAs")

for (tmp_PAtype in PA_types) {
  out_rds <- paste0(out_path,"result_2_conservation_effectiveness_since1963/Measure2_",tmp_PAtype,"_",suffix_name,"_break",my_break,".rds")
  if (!file.exists(out_rds)) {
    input_path <- paste0(out_path,"result_2_conservation_effectiveness_since1963/Measure2_",tmp_PAtype,"_",suffix_name,".rds")
    tmp_PAs <- readRDS(input_path) %>%
      select(matches("Year"))
    tmp_plot_df <- data.frame(
      prop = NA,
      year = NA, 
      count = NA
    )
    for (j in colnames(tmp_PAs)) {
      tmp_count <- NULL
      for (tmp_b in seq(0, 1, my_break)) {
        tmp_b_count <- sum(tmp_PAs[,j] >=tmp_b)
        tmp_count <- c(tmp_count, tmp_b_count)
      }
      tmp_plot_df_j <- data.frame(
        prop = seq(0, 1, my_break),
        year = rep(j, length(seq(0, 1, my_break))),
        count = tmp_count
      )
      tmp_plot_df <- rbind(tmp_plot_df, tmp_plot_df_j)
    }
    tmp_plot_df <- tmp_plot_df[-1,] %>% 
      setNames(c("Prop","Year", paste0(gsub("PAs","",tmp_PAtype),"count_",y))) %>%
      mutate(Year = gsub("Year_", "", Year),
             groups = paste0(Year,"_", Prop)) %>% 
      `rownames<-`(1:nrow(.))
    saveRDS(tmp_plot_df, out_rds)
  }
}

```

## 3.3 Merge different types of PAs for measure 2 into table within specific intervals

```{r merge_prop}

my_break <- 0.2
suffix_name <- "every_30years"
PA_types <- c("All_PAs", "Panda_PAs", "Plant_PAs", "Other_PAs")
 
out_rds <- paste0(out_path,"result_2_conservation_effectiveness_since1963/Measure2_",length(PA_types),"Types_PAs_",suffix_name,"_merged.rds")

if (!file.exists(out_rds)) {
  tmp_ff<- list.files(paste0(out_path, "result_2_conservation_effectiveness_since1963/"),
                      paste0("Measure2.*", suffix_name,"_break",my_break,".rds"), full.names = T)
  df <- readRDS(tmp_ff[1]) %>% 
    select(matches("groups"))
  for (tmp_PAtype in PA_types) {
    ff <- tmp_ff[grepl(tmp_PAtype,tmp_ff)]
    tmp_rds <- readRDS(ff) %>% 
      select(matches(c("groups","count"))) %>%
      rename_all(~ gsub("_count", "", .))  %>% 
      data.frame()
    
    df <- merge(df, tmp_rds, by="groups", all=T)
    
    if (tmp_PAtype == "Panda_PAs") {
      rows_0 <- which(grepl("_0$",df$groups))
      rows_1 <- which(grepl("_1$",df$groups))
      col_panda <- colnames(df)[grepl("Panda",colnames(df))]
      na_rows_0 <- rows_0[which(sum(df[rows_0,col_panda], na.rm = T)==0)]
      na_rows_1 <- rows_1[which(sum(df[rows_0,col_panda], na.rm = T)==0)]
      if (sum(na_rows_1-na_rows_0 ==rep(6-1, length(na_rows_0)))==length(na_rows_1)) {
        if (length(na_rows_0)!=0) {
          for (i in 1:length(na_rows_0)) {
            na_row_0_sel <- na_rows_0[i]
            na_row_1_sel <- na_rows_1[i]
            df[na_row_0_sel:na_row_1_sel,col_panda] <- df[(na_row_0_sel-6):(na_row_1_sel-6),col_panda]
          }
        }
      }
    }
    if (tmp_PAtype == "Plant_PAs") {
      rows_0 <- which(grepl("_0$",df$groups))
      rows_1 <- which(grepl("_1$",df$groups))
      col_plant <- colnames(df)[grepl("Plant",colnames(df))]
      na_rows_0 <- rows_0[which(sum(df[rows_0,col_plant], na.rm = T)==0)]
      na_rows_1 <- rows_1[which(sum(df[rows_0,col_plant], na.rm = T)==0)]
      if (sum(na_rows_1-na_rows_0 ==rep(6-1, length(na_rows_0)))==length(na_rows_1)) {
        if (length(na_rows_0)!=0) {
          for (i in 1:length(na_rows_0)) {
            na_row_0_sel <- na_rows_0[i]
            na_row_1_sel <- na_rows_1[i]
            df[na_row_0_sel:na_row_1_sel,col_plant] <- df[(na_row_0_sel-6):(na_row_1_sel-6),col_plant]
          }
        }
      }
    }
    if (tmp_PAtype == "Other_PAs") {
      rows_0 <- which(grepl("_0$",df$groups))
      rows_1 <- which(grepl("_1$",df$groups))
      col_other <- colnames(df)[grepl("Other",colnames(df))]
      na_rows_0 <- rows_0[which(sum(df[rows_0,col_other], na.rm = T)==0)]
      na_rows_1 <- rows_1[which(sum(df[rows_0,col_other], na.rm = T)==0)]
      if (sum(na_rows_1-na_rows_0 ==rep(6-1, length(na_rows_0)))==length(na_rows_1)) {
        if (length(na_rows_0)!=0) {
          for (i in 1:length(na_rows_0)) {
            na_row_0_sel <- na_rows_0[i]
            na_row_1_sel <- na_rows_1[i]
            df[na_row_0_sel:na_row_1_sel,col_other] <- df[(na_row_0_sel-6):(na_row_1_sel-6),col_other]
          }
        }
      }
    }
    saveRDS(df, out_rds)
  }
}
```

## 3.4 Prepare the datasets for plotting the effectiveness of measure 2

```{r Prop_for_plot}

my_break <- 0.2
suffix_name <- "every_30years"
PA_types <- c("All_PAs", "Panda_PAs", "Plant_PAs", "Other_PAs")

out_csv <- paste0(out_path,"result_2_conservation_effectiveness_since1963/Plot_Measure2_",length(PA_types),"Types_PAs_",suffix_name,"_break",my_break,".csv")
out_csv_melt <- gsub("\\.csv","\\_melt.csv",out_csv)
out_all <- c(out_csv, out_csv_melt)

if (sum(!file.exists(out_all)!=0)) {
  
  ff <- paste0(out_path,"result_2_conservation_effectiveness_since1963/Measure2_",length(PA_types),"Types_PAs_",suffix_name,"_merged.rds")
  
  tmp_df <- readRDS(ff)

  df_all <- tmp_df %>% 
    select(matches(c("groups","All"))) %>% 
    melt(id.vars = "groups", variable.name = "datasets", value.name = "All_count") %>% 
    mutate(prop = as.numeric(unlist(lapply(strsplit(groups, "_"), "[[", 2))),
           datasets = as.numeric(unlist(lapply(strsplit(as.character(datasets), "_"), "[[", 2))),
           year = paste0(unlist(lapply(strsplit(groups, "_"), "[[", 1)), "_",datasets)) %>% 
    select(matches(c("prop", "year","count")))
  
  df_panda <- tmp_df %>% 
    select(matches(c("groups","Panda"))) %>% 
    melt(id.vars = "groups", variable.name = "datasets", value.name = "Panda_count") %>% 
    mutate(prop = as.numeric(unlist(lapply(strsplit(groups, "_"), "[[", 2))),
           datasets = as.numeric(unlist(lapply(strsplit(as.character(datasets), "_"), "[[", 2))),
           year = paste0(unlist(lapply(strsplit(groups, "_"), "[[", 1)), "_",datasets)) %>% 
    select(matches(c("prop", "year","count")))
  
  df_plant <- tmp_df %>% 
    select(matches(c("groups","Plant"))) %>% 
    melt(id.vars = "groups", variable.name = "datasets", value.name = "Plant_count") %>% 
    mutate(prop = as.numeric(unlist(lapply(strsplit(groups, "_"), "[[", 2))),
           datasets = as.numeric(unlist(lapply(strsplit(as.character(datasets), "_"), "[[", 2))),
           year = paste0(unlist(lapply(strsplit(groups, "_"), "[[", 1)), "_",datasets)) %>% 
    select(matches(c("prop", "year","count")))
  
  df_other <- tmp_df %>% 
    select(matches(c("groups","Other"))) %>% 
    melt(id.vars = "groups", variable.name = "datasets", value.name = "Other_count") %>% 
    mutate(prop = as.numeric(unlist(lapply(strsplit(groups, "_"), "[[", 2))),
           datasets = as.numeric(unlist(lapply(strsplit(as.character(datasets), "_"), "[[", 2))),
           year = paste0(unlist(lapply(strsplit(groups, "_"), "[[", 1)), "_",datasets)) %>% 
    select(matches(c("prop", "year","count")))
  
  df_combine <- cbind(df_all, df_panda, df_plant, df_other) %>% 
    data.frame()
  
  
  if(table(as.character(df_combine$prop)==as.character(df_combine$prop.1))==nrow(df_combine)){df_combine$prop.1 <- NULL}
  if(table(as.character(df_combine$year)==as.character(df_combine$year.1))==nrow(df_combine)){df_combine$year.1 <- NULL}
  if(table(as.character(df_combine$prop)==as.character(df_combine$prop.2))==nrow(df_combine)){df_combine$prop.2 <- NULL}
  if(table(as.character(df_combine$year)==as.character(df_combine$year.2))==nrow(df_combine)){df_combine$year.2 <- NULL}
  if(table(as.character(df_combine$prop)==as.character(df_combine$prop.3))==nrow(df_combine)){df_combine$prop.3 <- NULL}
  if(table(as.character(df_combine$year)==as.character(df_combine$year.3))==nrow(df_combine)){df_combine$year.3 <- NULL}
  
  df_combine_melt <- df_combine %>% 
    mutate(groups= paste0(prop,"p_", year)) %>% 
    select(-c("prop", "year")) %>% 
    melt(id.vars = "groups", variable.name = "PAs", value.name = "count") %>% 
    mutate(PAs = gsub("_count","",PAs),
           prop = unlist(lapply(strsplit(groups, "p_"), "[[", 1)),
           year = paste0(PAs,"_",unlist(lapply(strsplit(groups, "p_"), "[[", 2)))) %>% 
    select(matches(c("prop", "year","count")))
  
  if(!file.exists(out_csv)) {
    write.csv(df_combine, out_csv)
  }
  if(!file.exists(out_csv_melt)) {
    write.csv(df_combine_melt, out_csv_melt)
  }
}

```
