---
title: "Analysis 8: Comparison of PAs effectiveness inoptimal, random and reality scenarios"
output: html_document
---

```{r setup, include=FALSE}

# set you working directory
setwd(my_wd <- 'path/to/your/data/dir/')
in_path <- paste0(my_wd,"input/")
out_path <- paste0(my_wd,"output/")


library(sf)
library(terra)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)

```

# 1 Optimal optimization without considering existing PAs and land-use change

```{r}

in_Rdata <- paste0(out_path, "result_4_conservation_optimiztion_our_method/prepared_datasets/source_datasets_optimization.RData")
load(in_Rdata);gc()

in_folder <- paste0(out_path, "result_8_effectiveness_comparison_optimal_random_reality/")

path_df3 <- path_df_strategy[1,1]
target <- 50
scenario <- "scenario0"
SSP_RCP <- "without_LUCC"
    
out_folder <- paste0(in_folder, scenario, "_target", target, "_",SSP_RCP,"//")
out_final <- paste0(out_folder, "final_results//")
out_process <- paste0(out_folder, "process_results//")  
if(!file.exists(out_folder)){dir.create(out_folder, recursive = TRUE)}
if(!file.exists(out_final)){dir.create(out_final, recursive = TRUE)}
if(!file.exists(out_process)) {dir.create(out_process, recursive = TRUE)}

if (length(list.files(out_final, "_final.rds$"))==0) {
  log_file <- file(paste0(out_folder, scenario, "_", SSP_RCP, "_", target, ".log"), "a")
  df3_PUs_mark <- readRDS(path_df3)
  tmp_values <- ifelse(df3_PUs_mark$richness==1, -1, NA) %>% setNames(scenario)
  
  if (SSP_RCP!="without_LUCC") {
    tmp_scenario <- df3_PUs_mark %>% select(matches(paste0("SSP")))
    tmp_values <- ifelse(tmp_scenario==1, -1, tmp_values)
  }
  df3_PUs_mark <- df3_PUs_mark%>% 
    mutate(scenario0 = as.numeric(tmp_values)) %>% 
    select(matches(paste0("SSP|",scenario))) %>% 
    setNames(c("LU_poten", "PUs_ID")) %>% 
    mutate(PUs_ID_new = PUs_ID)

  print(paste0("Start: ",scenario, "_", SSP_RCP, "_",target))
  cat(paste0("Start: ",scenario, "_", SSP_RCP, "_",target, "\n"), file = log_file, append = TRUE)
  
  thre_law_sp <- df2_spinfo$gbif_name[df2_spinfo$threat_sp==1| df2_spinfo$law_sp==1]
  priority_mat_occ <- as.numeric(df2_spinfo$priority_pixels) %>% 
    as.matrix()
  
  if (scenario!="scenario1"|SSP_RCP!="without_LUCC") {
    sel_PUs_unsu <- (df3_PUs_mark$PUs_ID==-1) & (!is.na(df3_PUs_mark$PUs_ID))
    unsu_PUs_sp_matrix <-  df1_PUs_sp[sel_PUs_unsu,]
    df2_spinfo$unsu_pixels <- colSums(unsu_PUs_sp_matrix)
    rm(unsu_PUs_sp_matrix)
  } else {
    df2_spinfo$unsu_pixels <- 0
  }
  
  
  if (length(list.files(out_process))!=0) {
    l <- 1
    if (length(list.files(out_process))==1) {
      i <- as.numeric(gsub(
        paste0("optimization_",scenario,"_",SSP_RCP,"_",target,"_Iteration|\\.rds"),
        "",list.files(out_process)))
    } else {
      i_first <- min(as.numeric(gsub(paste0("optimization_",scenario,"_",SSP_RCP,"_",target,"_Iteration|\\.rds"), "",list.files(out_process))))
      n <- ceiling(i_first/100)
      i <-  (length(list.files(out_process))-n)*100
    }
    rds_path <- list.files(out_process, paste0("Iteration",i,".rds"), full.names=T)
    v_PUs <- readRDS(rds_path)
    df3_PUs_mark$PUs_ID <- df3_PUs_mark$PUs_ID_new <- v_PUs
    sel_PUs <- is.na(df3_PUs_mark$PUs_ID_new)
    df_candidate_PUs <- df1_PUs_sp[sel_PUs,]
    df1_PUs_sp <- path_df_scenario <- NULL; gc()
  } else {
    l=2
    i=0
    sel_PUs <- is.na(df3_PUs_mark$PUs_ID_new)
    df_candidate_PUs <- df1_PUs_sp[sel_PUs,]
    df1_PUs_sp <- path_df_scenario <- NULL; gc()
  }
  
      

  df_PUs_scores <- data.frame(matrix(NA, nrow = nrow(df_candidate_PUs),ncol=4))
  colnames(df_PUs_scores) <- c("LU_weight","rank","PUs_score","PU_score")
  rownames(df_PUs_scores) <- rownames(df_candidate_PUs)
  df_PUs_scores$LU_weight <- (1-df3_PUs_mark$LU_poten[sel_PUs])
  gc()
  
  if (l==1) {
    sel_score <- is.na(df_PUs_scores$rank)
    unsu_sp <- df2_spinfo$gbif_name[df2_spinfo$unsu_pixels == df2_spinfo$total_pixels]
    covered_sp <- colnames(df_candidate_PUs)[!colnames(df_candidate_PUs)%in% unsu_sp]
    cols_covered_sp <- which(colnames(df_candidate_PUs) %in% covered_sp)
    mat_candidate_PUs_covered_sp  <- df_candidate_PUs[, cols_covered_sp] %>% as.matrix()
    mat_priority_PUs_covered_sp <- priority_mat_occ[cols_covered_sp] %>% as.matrix()
  
    df_PUs_scores$PUs_score <- as.numeric(mat_candidate_PUs_covered_sp %*% mat_priority_PUs_covered_sp);gc()

    mat_candidate_PUs_uc <- mat_candidate_PUs_covered_sp
    total_pixel_covered_sp <- df2_spinfo$total_pixels[covered_sp]
    for (c in 1:ncol(mat_candidate_PUs_uc)) {
      tmp_rate <- mat_candidate_PUs_uc[,c]/total_pixel_covered_sp[c]
      tmp_rate[tmp_rate<1] <- 0
      tmp_rate[is.na(tmp_rate)] <- 0
      mat_candidate_PUs_uc[,c] <- tmp_rate
      if(c==1){
        print("start protected more habitats")
        cat("start protected more habitats \n", file = log_file, append = TRUE)
      }
      cat(c, "\r")
    }
    df_PUs_scores$PUs_score_unique[sel_score] <- rowSums(mat_candidate_PUs_uc);gc()
    rm(mat_candidate_PUs_uc);gc()
  
    cover_tl_sp <- thre_law_sp[thre_law_sp %in% covered_sp]
    cols_cover_tl_sp <- which(colnames(df_candidate_PUs) %in% cover_tl_sp)
    mat_candidate_PUs_cover_tl  <- df_candidate_PUs[,cols_cover_tl_sp] %>% as.matrix()
    mat_priority_PUs_cover_tl  <- priority_mat_occ[cols_cover_tl_sp] %>% as.matrix()
    df_PUs_scores$PUs_score_tl[sel_score] <- (mat_candidate_PUs_cover_tl %*% mat_priority_PUs_cover_tl) %>% as.numeric()
    rm(mat_candidate_PUs_cover_tl);rm(mat_priority_PUs_cover_tl)
    rm(mat_candidate_PUs_covered_sp);rm(mat_priority_PUs_covered_sp);rm(priority_mat_occ)
    gap_sp <- NULL
    rm(df_candidate_PUs);gc()
  }
  gc()
  prot_pixel <- sum(df3_PUs_mark$PUs_ID!=-1,na.rm = T)
  target_pixel <- ceiling(nrow(df3_PUs_mark) * (target/100))
  number_loops <- ceiling((target_pixel-prot_pixel))    
  start.time <- Sys.time()
  
  for (l in 1:number_loops) {
    if (prot_pixel <= target_pixel) {
      
      # default setting
      loop.startime <- Sys.time()
      i <- i+1

      if (length(list.files(out_process))==0) {
        sel_score <- is.na(df_PUs_scores$rank)

        df_PUs_scores$PUs_score[!sel_score] <- df_PUs_scores$PU_score[!sel_score] <- 
          df_PUs_scores$PUs_score_unique[!sel_score] <- df_PUs_scores$PUs_score_tl[!sel_score] <- NA
        
        df2_spinfo$nonProt_pixels <- colSums(df_candidate_PUs)
        df2_spinfo$nonProt_unsu_pixels <- df2_spinfo$nonProt_pixels+df2_spinfo$unsu_pixels
        gap_sp <- df2_spinfo$gbif_name[df2_spinfo$nonProt_unsu_pixels == df2_spinfo$total_pixels]
        unsu_sp <- df2_spinfo$gbif_name[df2_spinfo$unsu_pixels == df2_spinfo$total_pixels]
        gap_sp <- gap_sp[!gap_sp%in% unsu_sp]
        cols_gap_sp <- which(colnames(df_candidate_PUs) %in% gap_sp)
    
        mat_candidate_PUs_gapsp <- df_candidate_PUs[,cols_gap_sp] %>% as.matrix()
        mat_priority_PUs_gapsp <- priority_mat_occ[cols_gap_sp] %>% as.matrix()
        df_PUs_scores$PUs_score[sel_score] <- as.numeric(mat_candidate_PUs_gapsp %*% mat_priority_PUs_gapsp);gc()
        
        mat_candidate_PUs_gu <- t(mat_candidate_PUs_gapsp)
        mat_candidate_PUs_gu <- t(mat_candidate_PUs_gu/df2_spinfo$total_pixels[gap_sp])
        mat_candidate_PUs_gu[mat_candidate_PUs_gu < 1] <- 0
        mat_candidate_PUs_gu[is.na(mat_candidate_PUs_gu)] <- 0
        
        df_PUs_scores$PUs_score_unique[sel_score] <- rowSums(mat_candidate_PUs_gu);gc() 
        rm(mat_candidate_PUs_gu)
        
        unpro_tl_sp <- thre_law_sp[thre_law_sp %in% gap_sp]
        cols_unpro_tl_sp <- which(colnames(df_candidate_PUs) %in% unpro_tl_sp)
        mat_candidate_PUs_gapsp_tl  <- df_candidate_PUs[,cols_unpro_tl_sp] %>% as.matrix()
        mat_priority_PUs_gapsp_tl  <- priority_mat_occ[cols_unpro_tl_sp] %>% as.matrix()
        df_PUs_scores$PUs_score_tl[sel_score] <- (mat_candidate_PUs_gapsp_tl %*% mat_priority_PUs_gapsp_tl) %>% as.numeric()
        
        if(length(gap_sp)==0){
          saveRDS(df3_PUs_mark$PUs_ID_new, paste0(out_process, "optimization_",scenario, "_", SSP_RCP, "_",target,"_Iteration",i-1, ".rds"))
        }          
        if (length(list.files(out_process))==1) {
          df_PUs_scores$PU_score <- NULL
          df_PUs_scores$PUs_score <- df_PUs_scores$PUs_score_unique <- df_PUs_scores$PUs_score_tl <- NA
          covered_sp <- colnames(df_candidate_PUs)[!colnames(df_candidate_PUs)%in% unsu_sp]
          cols_covered_sp <- which(colnames(df_candidate_PUs) %in% covered_sp)
          mat_candidate_PUs_covered_sp  <- df_candidate_PUs[, cols_covered_sp] %>% as.matrix()
          mat_priority_PUs_covered_sp <- priority_mat_occ[cols_covered_sp] %>% as.matrix()

          df_PUs_scores$PUs_score[sel_score] <- as.numeric(mat_candidate_PUs_covered_sp %*% mat_priority_PUs_covered_sp);gc()

          mat_candidate_PUs_uc <- mat_candidate_PUs_covered_sp
          total_pixel_covered_sp <- df2_spinfo$total_pixels[covered_sp]
          for (c in 1:ncol(mat_candidate_PUs_uc)) {
            tmp_rate <- mat_candidate_PUs_uc[,c]/total_pixel_covered_sp[c]
            tmp_rate[tmp_rate<1] <- 0
            tmp_rate[is.na(tmp_rate)] <- 0
            mat_candidate_PUs_uc[,c] <- tmp_rate
            if(c==1){
              print("start protected more habitats")
              cat("start protected more habitats \n", file = log_file, append = TRUE)
            }
            cat(c, "\r")
          }
          df_PUs_scores$PUs_score_unique[sel_score] <- rowSums(mat_candidate_PUs_uc);gc()
          rm(mat_candidate_PUs_uc)

          cover_tl_sp <- thre_law_sp[thre_law_sp %in% covered_sp]
          cols_cover_tl_sp <- which(colnames(df_candidate_PUs) %in% cover_tl_sp)
          mat_candidate_PUs_cover_tl  <- df_candidate_PUs[,cols_cover_tl_sp] %>% as.matrix()
          mat_priority_PUs_cover_tl  <- priority_mat_occ[cols_cover_tl_sp] %>% as.matrix()
          df_PUs_scores$PUs_score_tl[sel_score] <- (mat_candidate_PUs_cover_tl %*% mat_priority_PUs_cover_tl) %>% as.numeric()
          rm(mat_candidate_PUs_cover_tl);rm(mat_priority_PUs_cover_tl)
          rm(mat_candidate_PUs_covered_sp);rm(mat_priority_PUs_covered_sp);rm(priority_mat_occ)
          rm(df_candidate_PUs)
        }
      }
      
      
      if ((scenario!="scenario1") & (SSP_RCP=="all_SSPs_RCPs")) {
        df_PUs_scores$PUs_score <- df_PUs_scores$PUs_score * df_PUs_scores$LU_weight
        df_PUs_scores$PUs_score_unique <- df_PUs_scores$PUs_score_unique * df_PUs_scores$LU_weight
        df_PUs_scores$PUs_score_tl <- df_PUs_scores$PUs_score_tl * df_PUs_scores$LU_weight
      }
      
      sel_score <- is.na(df_PUs_scores$rank)   
      max_score <- max(df_PUs_scores$PUs_score[sel_score])
      sel_rows <- which(df_PUs_scores$PUs_score==max_score & is.na(df_PUs_scores$rank))
      
      if (length(sel_rows)==1) {
        sel_PU <- rownames(df_PUs_scores)[sel_rows]
      } else {
        max_unique <- max(df_PUs_scores$PUs_score_unique[sel_rows])
        ord_rows_unique <- which(df_PUs_scores$PUs_score_unique[sel_rows]==max_unique)
        sel_rows_unique <- sel_rows[ord_rows_unique]
        if (length(sel_rows_unique)==1) {
          sel_PU <- rownames(df_PUs_scores)[sel_rows_unique]
        } else {
          ord_rows_tl <- which.max(df_PUs_scores$PUs_score_tl[sel_rows])
          sel_rows_tl <- sel_rows[ord_rows_tl]
          sel_PU <- rownames(df_PUs_scores)[sel_rows_tl]
        }
      }
      
      df3_PUs_mark$PUs_ID_new[rownames(df3_PUs_mark)==sel_PU] <- i
      df_PUs_scores$rank[rownames(df_PUs_scores)==sel_PU] <- i
      if(length(gap_sp)>0){df_candidate_PUs <- df_candidate_PUs[rownames(df_candidate_PUs)!=sel_PU,]}
      prot_pixel <- sum(df3_PUs_mark$PUs_ID_new!=-1,na.rm = T)
      tmp_sum <- sum(df_PUs_scores$PUs_score[sel_score], na.rm = T)
      
      if (length(gap_sp)==0){
        if ((prot_pixel != target_pixel)){
          if (i %% 100 == 0) {
            df3_PUs_mark$PUs_ID_new[is.na(df3_PUs_mark$PUs_ID)] <- df_PUs_scores$rank
            saveRDS(df3_PUs_mark$PUs_ID_new, paste0(out_process, "optimization_",scenario, "_", SSP_RCP, "_",target,"_Iteration",i, ".rds"))
          }
        } else if (prot_pixel == target_pixel|tmp_sum == 0 ) {
          df3_PUs_mark$PUs_ID_new[is.na(df3_PUs_mark$PUs_ID)] <- df_PUs_scores$rank
          saveRDS(df3_PUs_mark$PUs_ID_new, paste0(out_final, "optimization_",scenario, "_", SSP_RCP, "_",target,"_Iteration",i, "_final.rds"))
        }
      }
      loop.endtime <- Sys.time()
      if(length(gap_sp)>0){
        cat("Iteration:",i,"unpro_sp:", length(gap_sp),"sum priority:", 
            tmp_sum, "Remaining:",target_pixel-prot_pixel, "=", prot_pixel,"/",target_pixel,
            "time:", loop.endtime-loop.startime, "total time:",loop.endtime-start.time, "scenario:", scenario, "target:", target, "SSP_RCP:",SSP_RCP,"dim:",dim(df_candidate_PUs), "\r")
        cat("Iteration:",i,"unpro_sp:", length(gap_sp),"sum priority:", 
            tmp_sum, "Remaining:",target_pixel-prot_pixel, "=", prot_pixel,"/",target_pixel,
            "time:", loop.endtime-loop.startime, "total time:",loop.endtime-start.time, "scenario:", scenario, "target:", target, "SSP_RCP:",SSP_RCP,"dim:",dim(df_candidate_PUs), "\n",  file = log_file, append = TRUE)
      } else {
        cat("Iteration:",i,"unpro_sp:", length(gap_sp),"sum priority:", tmp_sum , "Remaining:",target_pixel-prot_pixel, "=", prot_pixel,"/",target_pixel, "time:", loop.endtime-loop.startime, "total time:",loop.endtime-start.time, "scenario:", scenario, "target:", target, "SSP_RCP:",SSP_RCP, "\r")
        cat("Iteration:",i,"unpro_sp:", length(gap_sp),"sum priority:", tmp_sum , "Remaining:",target_pixel-prot_pixel, "=", prot_pixel,"/",target_pixel, "time:", loop.endtime-loop.startime, "total time:",loop.endtime-start.time, "scenario:", scenario, "target:", target, "SSP_RCP:",SSP_RCP, "\n",  file = log_file, append = TRUE)
      }
      if (tmp_sum==0){
        next
      }
      gc()
    }
  }
} else {
  print(paste0("Done: ",scenario, "_", SSP_RCP, "_",target))
}


```

# 2 Random optimization without considering existing PAs and land-use change (repeated 100 times)

```{r}

random_times <- 1:100
target <- 50
set.seed(20240513)

df_template <- readRDS(paste0(in_path, "PAs_coverage/pixels_covered_by_PAs.rds"))
out_folder <- paste0(out_path, "result_8_effectiveness_comparison_optimal_random_reality/Random_selection_target50_without_LUCC/")
df_v_scenarios <- v_scenario <- rep(NA, length(df_template))

if (length(list.files(out_folder))!=max(random_times)) {
  tmp_sample <- v_scenario[is.na(v_scenario)]
  No_target <- ceiling(length(df_v_scenarios) * (target/100))
  for (r_time in random_times) {
    out_rds <- paste0(out_folder, basename(out_folder),"_random_",r_time, ".rds")
    if (!file.exists(paste0(out_rds))) {
      my_each_random <- sample(length(tmp_sample))
      final_random <- ifelse(my_each_random<= No_target, my_each_random, NA)
      df_v_scenarios[is.na(v_scenario)] <- final_random 
      saveRDS(df_v_scenarios, out_rds)
      cat(gsub("\\.rds","",basename(out_rds)), "\r")
    }
  }
}

```

# 3 Prepare the datasets combining three optimization scenarios

scenario 1: Optimal optimization without considering existing PAs and land-use change
scenario 2: Realistic optimization incorporates both existing PAs and land-use change (=optimization strategy2)
scenario 3: Randomized optimization without considering land-use change (repeated 100 times)

```{r}

out_rds_binary <- paste0(out_path, "result_8_effectiveness_comparison_optimal_random_reality/effectiveness_comparison_optimal_random_reality_binary.rds")
out_rds <- paste0(out_path, "result_8_effectiveness_comparison_optimal_random_reality/effectiveness_comparison_optimal_random_reality.rds")
out_all <- c(out_rds_binary, out_rds)

in_PUs_sp_binary <- paste0(in_path, "pixels_by_species/pixels_by_species_binary_AOH30m_aggregated.rds")
in_PUs_sp <- paste0(in_path, "pixels_by_species/pixels_by_species_AOH30m_aggregated.rds")

if (sum(!file.exists(out_all))!=0) {
  
  folder_reality <- paste0(out_path,"result_4_conservation_optimiztion_our_method/")
  path_reality <- list.files(list.files(list.files(folder_reality, "2_target50.*", full.names = T), 
                                          "final_results$", full.names = T), ".rds$", full.names = T)
  folder_optimal <- paste0(out_path, "result_8_effectiveness_comparison_optimal_random_reality/scenario0_target50_without_LUCC/")
  path_optimal <- list.files(paste0(folder_optimal, "final_results"), ".rds", full.names = T)
  
  folder_random <-  paste0(out_path, "result_8_effectiveness_comparison_optimal_random_reality/Random_selection_target50_without_LUCC/")
  path_random <- list.files(folder_random, full.names = T)
  names_random <- gsub("Random_selection_target50_without_LUCC_|\\.rds",
                            "",basename(path_random))
  
  path_all <- c(path_reality, path_optimal,path_random)
  name_all <- c("reality", "optimal",names_random)
   
  df_all <- readRDS(path_all[1])
  df_all <- matrix(NA, nrow=length(df_all))
  for (i in 1:length(path_all)) {
    tmp_path <- path_all[i]
    tmp_name <- name_all[i]
    tmp_df <- readRDS(tmp_path) %>% 
      data.frame() %>% 
      setNames(tmp_name)
    df_all <- cbind(df_all, tmp_df)
  }
  if (sum(!file.exists(out_rds_binary))!=0) {
    df_PUs_sp_binary <- readRDS(in_PUs_sp_binary)
    df_all_binary <- df_all %>% 
      select(-df_all) %>% 
      cbind(., df_PUs_sp_binary)
    saveRDS(df_all_binary, out_rds_binary)
    rm(df_PUs_sp_binary);rm(df_all_binary)
  }
  if (sum(!file.exists(out_rds))!=0) {
    df_PUs_sp <- readRDS(in_PUs_sp)
    df_all <- df_all %>% 
      select(-df_all) %>% 
      cbind(., df_PUs_sp)
    saveRDS(df_all, out_rds)
    rm(df_PUs_sp);rm(df_all)
  }
}
      
```

# 4 Prepare the datasets to compare the effectiveness in covered species under three scenarios

```{r}

gbif_name <- read.csv(paste0(in_path, "species_info/species_list.csv"),row = 1) %>% 
  pull(gbif_name);gc()
in_rds <- paste0(out_path, "result_8_effectiveness_comparison_optimal_random_reality/effectiveness_comparison_optimal_random_reality_binary.rds")
df_sp <- readRDS(in_rds);gc()
df_info <- df_sp %>% 
  select(-all_of(gbif_name))
df_sp <- df_sp %>% select(all_of(gbif_name));gc()
df_out <- data.frame(prop_area = seq(0,60,by=0.01))

out_rds <- paste0(out_path, "result_8_effectiveness_comparison_optimal_random_reality/effectiveness_number_covered_species_optimal_random_reality.rds")
Total_counts <- nrow(df_sp)

if(!file.exists(out_rds)) {
  all_names <- colnames(df_info)
  for (tmp_n in 1:length(all_names)) {
    tmp_col <- all_names[tmp_n]
    tmp_df <- df_info %>%
      select(all_of(tmp_col)) %>% 
      setNames("scenario") %>% 
      cbind(.,df_sp) %>% 
      filter(!scenario==-1&!is.na(scenario)) %>% 
      arrange(scenario) %>% 
      mutate(rank_sce = rank(scenario, ties.method = "max"),
             prop_area = round(rank_sce/nrow(df_sp)*100,2)) %>% 
      select(prop_area,everything(),-rank_sce,-scenario) %>%
      group_by(prop_area) %>% 
      summarize(across(everything(),~sum(.x, na.rm = TRUE)))
    
    gc()
    
    tmp_prop_area <- tmp_df %>% select(prop_area)
    tmp_df <- tmp_df %>% select(-prop_area) 
    max_richness <- sum(colSums(tmp_df,na.rm=T)!=0)
    richness <- c()
    for(tmp_r in 1:nrow(tmp_df)){
      tmp_df_sp <- tmp_df[1:tmp_r,]
      tmp_colsum <- colSums(tmp_df_sp,na.rm=T)!=0;rm(tmp_df_sp)
      tmp_richness <- sum(tmp_colsum)
      richness <- c(richness, tmp_richness)
      gc()
      cat(paste0(tmp_n,"/", length(all_names),
                 "; - ",tmp_r," - ", nrow(tmp_df), 
                 "; ",tmp_richness, "- ",max_richness,
                 "; scenario: ", tmp_col,"\r"))
      if(tmp_richness==max_richness){
        break
      }
    }
    gc();rm(tmp_df)
    df_prop_area <- data.frame(prop_area=tmp_prop_area$prop_area[1:length(richness)],
                           richness=richness) %>% 
      merge(.,tmp_prop_area,all.y=T) %>% 
      mutate(richness=ifelse(is.na(richness), max(richness,na.rm = T), richness)) %>% 
      setNames(c("prop_area", tmp_col))
    gc();rm(tmp_prop_area)
    df_out <- 
      merge(df_out,df_prop_area,by="prop_area",all.x=T) 
  }
  saveRDS(df_out, out_rds)
}

```

# 5 Prepare the datasets to compare the effectiveness in covered habitat under three scenarios

```{r}

in_rds <-  paste0(out_path, "result_8_effectiveness_comparison_optimal_random_reality/effectiveness_comparison_optimal_random_reality.rds")
in_PUs_sp <-  paste0(in_path, "pixels_by_species/pixels_by_species_AOH30m_aggregated.rds")
my_break <- 0.2
out_rds <- paste0(out_path, "result_8_effectiveness_comparison_optimal_random_reality/effectiveness_prop_covered_habitat_optimal_random_reality_freq_break",my_break,".rds")

if (!file.exists(out_rds)) {
  
  ras_PAs <- rast(paste0(in_path, "PAs_coverage/PAs_ID_dynamic.tif"))
  shp_PAs <- st_read(paste0(in_path, "PAs_coverage/PAs_categories_info.shp")) %>% 
    st_drop_geometry(.) %>% 
    select(OBJECTID, Year) %>% 
    setNames(c("ID_PAs", "Year"))
  
  v_PAs <- values(ras_PAs) %>% 
    data.frame() %>% 
    filter(!is.na(.)) %>% 
    setNames("ID_PAs") %>% 
    mutate(ID=1:nrow(.),
           No_pixel =1,
           ID_PAs = ifelse(ID_PAs<200, ID_PAs, ID_PAs-200)) %>% 
    merge(., shp_PAs, by="ID_PAs", all=T) %>% 
    arrange(Year) %>% 
    mutate(rank=ifelse(!is.na(Year),1:length(!is.na(Year)),NA)) %>% 
    arrange(ID)
  
  df_PAs <- values(ras_PAs) %>% 
    data.frame() %>% 
    filter(!is.na(.)) %>% 
    setNames("ID_PAs") %>% 
    mutate(No_pixel =1,
           ID_PAs = ifelse(ID_PAs<200, ID_PAs, ID_PAs-200)) %>% 
    group_by(ID_PAs) %>% 
    summarize(across(everything(), sum)) %>% 
    filter(ID_PAs!=0) %>%  
    merge(., shp_PAs, by="ID_PAs", all=T) %>% 
    group_by(Year) %>% 
    summarize(across(No_pixel, sum)) %>%
    mutate(Sum_pixel=cumsum(No_pixel))
  
  df_sel <- df_PAs %>% 
    filter(grepl("1963|1983|2003|2021",Year)) %>% 
    pull(Sum_pixel) %>% 
    c(.,ceiling(nrow(v_PAs)*0.3),
               ceiling(nrow(v_PAs)*0.5))
  
  gbif_name <- read.csv(paste0(in_path, "species_info/species_list.csv"),row = 1) %>% 
    pull(gbif_name)
  df_sp <- readRDS(in_rds);gc()
  
  df_info <- df_sp %>% 
    select(-all_of(gbif_name))
  
  my_scenario <- c("year_1963","year_1983", "year_2003",
                   "year_2023","year_2030", "year_2050")
  
  which_rows <- which(df_info$reality>0)
  v_scenario2 <- df_info$reality[which_rows]+max(v_PAs$rank,na.rm = T)
  df_info$reality[which_rows] <- v_scenario2
  which_rows <- which(df_info$reality==0)
  df_info$reality[which_rows] <- v_PAs$rank[which_rows]
  
  sel_NA <- intersect(which(df_sp$optimal==-1),
                      which(df_sp$reality==-1));gc()
  
  if (length(sel_NA)!=0) {
    df_sp <- df_sp[-sel_NA,]
    df_info <- df_info[-sel_NA,]
  }
  
  df_sp <- df_sp %>% select(all_of(gbif_name));gc()
  df_stat <- data.frame(gbif_name = gbif_name,
                       SC_pixel = colSums(df_sp, na.rm = T));gc()
  rownames(df_stat) <- NULL
  
  all_names <- colnames(df_info)
  for (tmp_n in 1:length(all_names)) {
    tmp_col <- all_names[tmp_n]
    tmp_df <- df_info %>%
      select(all_of(tmp_col)) %>% 
      setNames("scenario") %>% 
      cbind(.,df_sp) %>% 
      filter(!scenario==-1&!is.na(scenario)) %>% 
      arrange(scenario)
    my_sel <- df_sel
    for (tmp_s in 1:length(my_sel)) {
      tmp_sel <- my_sel[tmp_s]
      tmp_scenario <- my_scenario[tmp_s]
      tmp_sel_df <- tmp_df %>% 
        filter(scenario<=tmp_sel) %>% 
        select(-scenario)
      tmp_colsums <- colSums(tmp_sel_df, na.rm = T)
      df_stat <- df_stat %>% 
        mutate(!!paste0(tmp_col,"_",tmp_scenario):=tmp_colsums);gc()
      cat(paste0(tmp_n,"/", length(all_names),
                 "; - ",tmp_s," - ", length(my_sel),
                 "; PAs scenario: ", "- ",tmp_scenario,
                 "; scenario: ", tmp_col,"\r"))
    }
  }
  
  df_stat_info <- df_stat %>% 
    select(gbif_name,SC_pixel)
  df_stat_prop <- df_stat %>% 
    select(-gbif_name,-SC_pixel)/df_stat_info$SC_pixel
  df_stat_prop <- cbind(df_stat_info,df_stat_prop)
  
  tmp_df_sp <- df_stat_prop %>% 
      select(-gbif_name,-SC_pixel)
    
  tmp_plot_df <- data.frame(prop = NA, group = NA, count = NA)

  for (j in colnames(tmp_df_sp)) {
    tmp_count <- NULL
    for (tmp_b in seq(0, 1, my_break)) {
      tmp_b_count <- sum(tmp_df_sp[,j] >=tmp_b)
      tmp_count <- c(tmp_count, tmp_b_count)
    }
    tmp_plot_df_j <- data.frame(
      prop = seq(0, 1, my_break),
      group = rep(j, length(seq(0, 1, my_break))),
      count = tmp_count)
    tmp_plot_df <- rbind(tmp_plot_df, tmp_plot_df_j)
  }
      
  tmp_plot_df <- tmp_plot_df[-1,] %>% 
    setNames(c("Prop","group", "count")) %>%          
    `rownames<-`(1:nrow(.)) 
  
  saveRDS(tmp_plot_df, out_rds)
}

```